\section{Introduction}

The rapid proliferation of smartphones represents both a challenge and an
opportunity for wireless spectrum allocation. On one hand, smartphones are
competing for the same limited spectrum already crowded with other devices.
On the other hand, the growing network of mobile, always-on smartphones
constitutes an unprecedented new source of detailed measurements of spectrum
usage and availability. In addition, because smartphones are always on but
idle much of the time, they are in a unique position to provide measurements
about \textit{other} nearby active wireless devices---such as laptops,
tablets, or other smartphones---meaning that measurements can be performed
passively without disturbing devices actively using the spectrum, a
capability that has not been exploited by other systems designed to
incorporate client-side feedback into spectrum allocation. This type of
measurement also improves on existing approaches that use periodic planned
site surveys, since the data they provide is neither continuous nor
representative of the experience of actual network users.

As an example of the type of spectrum adaptation smartphones can enable, as a
user engages with their laptop their smartphone is nearby and capable of
helping the laptop improve its usage of the spectrum by scanning other
channels and determining whether both the client and access point should move
to a different, less-congested portion of the spectrum. Measurements from
passive smartphones can also improve the wireless performance of other users:
as a second example, as one user engages with their smartphone, the
smartphone of a second user sitting nearby may be inactive and able and
willing to perform measurements. Smartphones can also provide continuous
observations of wireless performance to help providers provision and maintain
their networks. We refer to these approaches collectively as
\textbf{c}rowdsourcing \textbf{a}ccess \textbf{n}etwork \textbf{s}pectrum
\textbf{a}llocation using \textbf{s}martphones, or \textbf{CANSAS}.

\sloppypar{Realizing CANSAS requires novel integration between smartphones, an adaptive
network, and algorithms enabling cooperative spectrum allocation on both
short and long timescales. This paper describes a prototype system
implementing CANSAS for Wifi networks called \PS{}. \PS{} uses a
publish-subscribe architecture to perform both asynchronous and synchronous
measurement collection from passive smartphones and use collected data to
improve the performance of active clients. Measurements are used as inputs to
new algorithms---including game-theoretic approaches---that perform joint
channel assignment, power control, rate selection, and transmission
scheduling to improve spectrum efficiency assuming a variety of different
network structures, including both fully-cooperative settings and cases where
multiple independently-administered networks compete for the local spectrum
resources. These scenarios are representative of both typical home and
corporate Wifi deployments. \PS{} measurements are also provided to network
administrators in order to perform long-term monitoring and network
maintenance. \PS{} provides a platform for testing these new
algorithms and determining how to incentivize client measurement collection,
assuring that collected measurements are accurate, and minimizing the energy
overhead for participating devices.}PocketSniffer 

The rest of this paper is structured as follows. We begin by describing how
using passive smartphones improves on previous client-side measurement
approaches in Section~\ref{sec-related}. Section~\ref{sec-design} briefly
describes the design of the data collection portions of \PS{} while
Section~\ref{sec-algorithms} discusses several game-theoretic algorithms that
utilize \PS{} measurements to perform spectrum adaptation. We
continue by presenting results from a prototype \PS{} system in
Section~\ref{sec-results} before discussing open challenges in
Section~\ref{sec-challenges} and concluding in Section~\ref{sec-conclusions}.
